---
title: "Datathon"
output: pdf_document
date: "2023-02-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(stringr)

#loading data
Dcrime <- read.csv("Crime.csv", header = TRUE)
DcrimeOriginal <- Dcrime #original without any cleaning


# Checking Empty date for Start, end, report date of crime
EmptyEnd <- Dcrime %>% filter(Offense.End.DateTime == "") #452006 obs
EmptyStart <- Dcrime %>% filter(Offense.Start.DateTime == "") #1203 obs
EmptyReport <- Dcrime %>% filter(Report.DateTime == "") #0 obs
# we will use Report.DateTime for Date

#selecting necessary columns
colnames(Dcrime)
colnames(Dcrime[5:15])
Dcrime <- Dcrime[5:15]

# Only for 2022 Offense.Start.DateTime
Dcrime22 <- Dcrime %>% filter(str_detect(Report.DateTime, "2022"))

table(Dcrime22$Crime.Against.Category)
table(Dcrime22$Offense)

table(Dcrime22$Offense.Parent.Group)
names(table(Dcrime22$Offense.Parent.Group)) #31 -- we will use this.
names(table(Dcrime$Offense)) #59

#looking at areas if we have NA MCPP
table(Dcrime22$MCPP)
```

One way to cluster the crimes into three different groups based on how serious they are could be:

Cluster 1: Least serious crimes

CURFEW/LOITERING/VAGRANCY VIOLATIONS
DRUNKENNESS
FAMILY OFFENSES, NONVIOLENT
FRAUD OFFENSES
GAMBLING OFFENSES
LIQUOR LAW VIOLATIONS
PORNOGRAPHY/OBSCENE MATERIAL
PROSTITUTION OFFENSES
TRESPASS OF REAL PROPERTY


Cluster 2: Moderately serious crimes

ANIMAL CRUELTY
BAD CHECKS
COUNTERFEITING/FORGERY
DESTRUCTION/DAMAGE/VANDALISM OF PROPERTY
EMBEZZLEMENT
KIDNAPPING/ABDUCTION
LARCENY-THEFT
STOLEN PROPERTY OFFENSES


Cluster 3: Most serious crimes

ARSON
ASSAULT OFFENSES
BURGLARY/BREAKING&ENTERING
DRIVING UNDER THE INFLUENCE
DRUG/NARCOTIC OFFENSES
EXTORTION/BLACKMAIL
HOMICIDE OFFENSES
HUMAN TRAFFICKING
MOTOR VEHICLE THEFT
PEEPING TOM
ROBBERY
SEX OFFENSES
SEX OFFENSES, CONSENSUAL
WEAPON LAW VIOLATIONS

Of course, this is just one possible way to cluster the crimes and other criteria or methods could be used depending on the context and the purpose of the analysis.


```{r}

Group22 <- Dcrime22 %>% 
  group_by(MCPP, Offense.Parent.Group) %>% 
  summarize(n = n(), .groups = "drop")

Group22 <- slice(Group22, -1) #Removing NA MCPP

head(Group22)


library(tidyr)
Matix22 <- pivot_wider(Group22, names_from = Offense.Parent.Group, values_from = n, values_fill = 0) # looking at all crime numbers (column) * MCPP (row)

Matix22 <- Matix22[-c(9, 59), ]
 #Removing CAP HILL 9th row, UNKNOWN 59th row


# Cluster 1: Least serious crimes
cols_least <- c("CURFEW/LOITERING/VAGRANCY VIOLATIONS", "DRUNKENNESS", "FAMILY OFFENSES, NONVIOLENT", "FRAUD OFFENSES", "GAMBLING OFFENSES", "LIQUOR LAW VIOLATIONS", "PORNOGRAPHY/OBSCENE MATERIAL", "PROSTITUTION OFFENSES", "TRESPASS OF REAL PROPERTY")

cols_least_n <- grep(paste(cols_least, collapse="|"), colnames(Matix22)) 
Least <- Matix22[cols_least_n]

# binding column for area name
Least <- cbind(Matix22$MCPP, Least)
colnames(Least)[1] <- "MCPP"
write.csv(Least, "Least.csv")


# Cluster 2: Moderately serious crimes
cols_moderate <- c("ANIMAL CRUELTY", "BAD CHECKS", "COUNTERFEITING/FORGERY", "DESTRUCTION/DAMAGE/VANDALISM OF PROPERTY", "EMBEZZLEMENT", "KIDNAPPING/ABDUCTION", "LARCENY-THEFT", "STOLEN PROPERTY OFFENSES")

cols_moderate_n <- grep(paste(cols_moderate, collapse="|"), colnames(Matix22)) 
Moderate <- Matix22[cols_moderate_n]
# binding column for area name
Moderate <- cbind(Matix22$MCPP, Moderate)
colnames(Moderate)[1] <- "MCPP"
write.csv(Moderate, "Moderate.csv")

# Cluster 3: Most  serious crimes
cols_most <- c("ARSON", "ASSAULT OFFENSES", "BURGLARY/BREAKING&ENTERING", "DRIVING UNDER THE INFLUENCE", "DRUG/NARCOTIC OFFENSES", "EXTORTION/BLACKMAIL", "HOMICIDE OFFENSES", "HUMAN TRAFFICKING", "MOTOR VEHICLE THEFT", "PEEPING TOM", "ROBBERY", "SEX OFFENSES", "SEX OFFENSES, CONSENSUAL", "WEAPON LAW VIOLATIONS")

cols_most_n <- grep(paste(cols_most, collapse="|"), colnames(Matix22)) 
Most <- Matix22[cols_most_n]
# binding column for area name
Most <- cbind(Matix22$MCPP, Most)
colnames(Most)[1] <- "MCPP"
write.csv(Most, "Most.csv")





my_shapefile <- arrange(my_shapefile, NEIGHBORHO)




Matix22 <- Matix22 %>%
  mutate(Serious == ifelse())


# Set row names to MCPP column
row.names(new_df) <- new_df$MCPP


Unknown <- Dcrime %>% filter(MCPP == "UNKNOWN")
table(Dcrime$Beat)

U1 <- Dcrime %>% filter(Beat == "U1")

library(sf)
my_shapefile <- st_read("Micro_Community_Policing_Plans.shp")
plot(my_shapefile)
d1 <- as.list(my_shapefile$NEIGHBORHO)
d2 <- as.list(Dcrime$MCPP)
do.call(paste0, d1) %in% do.call(paste0, d2)



max(Dcrime$Report.DateTime)

```